---
jupyter:
  jupytext:
    formats: ipynb,Rmd,R
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.9.1
  kernelspec:
    display_name: R
    language: R
    name: ir
---

We can write notebooks using a couple different IDEs, popular being 

1. [jupyter](https://jupyter.org/install) which is a bit more advanced and requires installing [anaconda](https://www.anaconda.com/)

2. [rstudio](https://rstudio.com/products/rstudio/)

either can be used to write code in [R](https://www.r-project.org/)

```{r}
print('hello world')
```

We can use R as a calculator

```{r}
1+1
```

```{r}
1+2*5
```

there are also some built-in constants in R

```{r}
pi
exp(1)
```

there are also special infinite values

```{r}
1/0
-1/0
```

```{r}
0/0
```

to denote nothing we have

```{r}
NULL
```

to denote a missing value

```{r}
NA
```

# Variable assignment and simple objects

```{r}
x = 1
y <- 2
3 -> z
x;y;z
print(x)
```

# Vectors

```{r}
x = c(5,3,7)
x
```

```{r}
x[1]
```

```{r}
x[c(-1,-2)]
```

we can make a vector out of anything

```{r}
truths <- c(TRUE, FALSE, TRUE)
```

```{r}
x[truths]
```

we can make a matrix using the ``matrix`` function

```{r}
1:25
X <- matrix(1:25,nrow=5,byrow=TRUE)
X
```

```{r}
Y <- array(25:49,c(5,5))
Y
```

matrix multiplication is done with the operator ``%*%``

```{r}
X %*% Y
```

# Flow Control

```{r}
A = FALSE
if(A){
    print("TRUE")
} else{
    print("FALSE")
}
```

```{r}
for(i in 1:10){
    print(i)
}
```

```{r}
i = 1
while( i <= 10){
    print(i)
    i = i + 1
}
```

there is a good tutorial by Dr. Leemis [here](http://www.math.wm.edu/~leemis/r.pdf)
